Deep Reinforcement Learning for Conservation Decisions

### 2021-05-10 manuscript/manuscript.Rmd
short paragraph on ethics/power in discussion

Any application of RL can be divided into two components: an *environment* and an *agent*.
An *environment* is typically a computer simulation, though it is possible to use the real world as the RL environment [@robotics].
Over the course of an episode, an *agent*, which is often a computer program, continuously interacts with the environment. 
At each time step, the agent performs an *action*, and as a result of this action, the environment will determine the next *state* and numerical *reward* that the agent will observe.
The goal of the agent is to learn how to maximize its expected cumulative reward.
The agent learns how to achieve this objective during a period called *training*.
In training, the agent *explores* the available states and actions. 
Once the agent comes across a highly rewarding sequence of states and actions, the agent will reinforce this behavior so that it is more likely for the agent to *exploit* the same high reward trajectory in future episodes.
Throughout this process, the agent's behavior is codified into what is called a *policy*, which describes what action an agent should take for a given state.

To solve the RL problem, algorithms either take a *model-free* or *model-based* approach. 
The distinction is that *model-free* algorithms do not use or attempt to learn a model of the environment.
Conversely, *Model-based* algorithms either learn or use a model of the environment to achieve the RL objective.
A trade-off between these approaches is that when it is possible to quickly learn a model of the environment or the model is already known, model-based algorithms tend to require much less interaction with the environment to learn good-performing policies [@mbpo].
Yet, frequently, learning a model of the environment is very difficult, and in these cases, model-free algorithms tend to outperform [@CITE].
Since most of the recent focus in RL has gone towards model-free algorithms, we will focus on model-free agents for the rest of the paper.

Among model-free algorithms, there is divergence around what functions the agent is attempting to learn in order to achieve the RL objective.
Generally, model-free agents are either learning a policy function, a value function or both.
For context, *value functions* are proxies for how high of a cumulative reward an agent can expect to receive from a given state or state-action pair.
If an agent knows the value function, then the agent can find the optimal action at any state by selecting the action that will maximize the value function in expectation.
The class of algorithms that exclusively try to learn a value function are called *value learning* methods.
*Policy gradient* algorithms exclusively learn the policy.
And, lastly, *Actor-critic* algorithms attempt to learn both a policy function and a value function, whereby the value function, which is learned separately from the policy, is used to inform the agent on the goodness of a selected action.
There is a handful of advantages and disadvantages for each of these approaches that we will not go in depth on -- for more, see [@CITE]; yet, the main point of distinction comes from the bias-variance trade-off.
Policy gradients tend to struggle with high variance, value-learning tends to be biased while actor-critic methods balance these two concerns.

Neural networks become useful in RL when the environment has a large state-action space, which happens frequently with realistic decision-making problems. 
Classic RL algorithms, like policy iteration, value iteration and Actor-Critic [@suttonbarto], fail when the state-action space is large, e.g. when the possible states and actions can not be represented in a tractable table [@CITE].
Yet, over the last 10 years, researchers have shown that neural network-based RL algorithms can perform well on previously unsolvable environments.
Neural networks have been widely useful in reinforcement learning because neural nets have the property of being universal function approximators [@CITE].
Whenever there is a need for an agent to approximate some function, say a policy function, value function or a transition function, neural networks can be used in this capacity.
While there are other function approximators that can be used in RL, e.g. Gaussian processes, neural networks have excelled in this role because of their ability to learn complex, non-linear, high dimensional functions and their ability to adapt given new information [@CITE].
We will not discuss in detail how neural networks work -- see [@CITE] for further background --, but the basic premise is that at each learning step, the agent, say a policy gradient-based agent, will adjust the parameters in its policy neural network so that according to past experiences, the agent will be more likely to engage in highly rewarding behavior in the future.

@@

<!-- 
Placeholder table as figure, rather hacky (pandoc doesn't like `longtable`) 
Make a nice table box in latex instead if we're keeping this
-->
\begin{figure}
\centering
\includegraphics[width=0.9\linewidth]{table1.pdf}
\end{figure}

@@

>>>>

<!-- something to introduce this section/transition from the intro might be helpful: "We start with a brief overview of RL problems and terminology:" -->

Any application of RL can be divided into two components: an *environment* and an *agent*.
An *environment* is typically a computer simulation, though it is possible to use the real world as the RL environment [@robotics].
Over the course of an episode, an *agent*, which is often a computer program, continuously interacts with the environment. 
At each time step, the agent performs an *action*, and as a result of this action, the environment will determine the next *state* and numerical *reward* that the agent will observe.
The goal of the agent is to learn how to maximize its expected cumulative reward.
The agent learns how to achieve this objective during a period called *training*.
In training, the agent *explores* the available states and actions. 
Once the agent comes across a highly rewarding sequence of states and actions, the agent will reinforce this behavior so that it is more likely for the agent to *exploit* the same high reward trajectory in future episodes.
Throughout this process, the agent's behavior is codified into what is called a *policy*, which is a mapping of states to actions.

<!-- ^ this is great! I do think keeping a table of these terms would be helpful (and refering to that table in the text here) -->

To solve the RL problem, algorithms either take a *model-free* or *model-based* approach. 
The distinction is that *model-free* algorithms do not have access to a model of the environment; yet, *model-based* algorithms use a model of the environment to achieve the RL objective.
A trade-off between these approaches is that when it is possible to quickly learn a model of the environment or the model is already known, model-based algorithms tend to require much less interaction with the environment to learn good-performing policies [@mbpo].
Yet, frequently, learning a model of the environment is very difficult, and in these cases, model-free algorithms tend to outperform [@CITE].
Since most of the recent focus in RL has gone towards model-free algorithms, we will focus on model-free agents for the rest of the paper.

Amongst model-free algorithms, there is divergence around what functions the agent is attempting to learn.
Generally, model-free agents are either learning a policy function, a value function or both.
For context, *value functions* are proxies for how high of a cumulative reward an agent can expect to receive from a given state or state-action pair.
If an agent knows the value function, then the agent can find the optimal action at any state by selecting the action that will maximize the value function in expectation.
The class of algorithms that exclusively try to learn a value function are called *value learning* methods.
By contrast, *Policy gradient* algorithms exclusively learn the policy.
*Actor-critic* algorithms attempt to learn both a policy function and a value function, whereby the value function, which is learned separately from the policy, is used to inform the agent on the goodness of a selected action.
There is a handful of advantages and disadvantages for each of these approaches that we will not go in depth -- for more, see [@CITE]; yet, the main point of distinction comes from the bias-variance trade-off.
Policy gradients tend to struggle with high variance, value-learning tends to be biased while actor-critic methods balance these two concerns.

@@

<!-- 
Placeholder table as figure, rather hacky (pandoc doesn't like `longtable`) 
Make a nice table box in latex instead if we're keeping this
-->

<!-- maybe consider some combined conceptual figure/table?  But overall I really like the table and think it will be helpful for readers -->

\begin{figure}
\centering
\includegraphics[width=0.9\linewidth]{table1.pdf}
\end{figure}

@@

<!-- first pass at this -->

Critical engagement in the ethical and power issues presented by decision making algorithms will be imperative to ensuring that applications of RL to ecological challenges are fair and just. While widespread discussion of the ethics of decision making algorithms across sectors such as policing, finance, and employment, are a commonplace little has been discussed on the biases and implications of implementing these methods in the environmental context. If algorithms are to play a meaningful role in the allocation and exploitation of natural resources from fish to forests, good intentions [@Joppa] may not be sufficient to ensure equitable outcomes. Mitigation of the power asymmetries that might arise from algorithmic applications are key [cite caleb's paper?]. Engagement with experts in governance and ethics to ensure that applications of AI to ecological challenges are applied responsibly and transparently.
